Performance for the masses: Experiments with A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Section 3.2 also reviewed

We are also interested in measuring the scaling properties in this
model, that is, the relationship between the number of IPs
participating in an experiment and the time it takes to complete
it. This is shown in Figure \ref{fig:ipstime:w2}, which displays every
experiment in terms of time (in milliseconds) vs. the number of workers participating
in it. Although there is not a clear trend, the graph seems to say
that the time needed for finding the solution in this evolutionary
algorithm does not depend on the number of workers participating in
it. This does not mean that it is independent on the number of {\em
  simultaneous} users, which is probably the case although it is more
difficult to measure. There seems to be a trend towards a decreasing
standard deviation in the time, with more volunteers adding both {\em
  robustness} to an experiment, and certainty in the time it is taking. 
 However, more experiments would be needed to check this hypothesis.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.49\linewidth]{gev-fit-ww.png}
\includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
\caption{Ranked number of PUTs per worker and GEV fit (in light
  or blue color, left) 
and Weibull (in lighter or blue shade, right) for the {\sf NodIO-W$^2$} model.}  
\label{fig:gev:w2}
% File ../data/ips-puts.R
\end{figure}
%
As we did in the previous subsection, we have also fitted the number
of HTTP PUTs per worker to a  Generalized Extreme Value (GEV) function, with the result shown
in Table \ref{tab:puts:ww}. Comparing this table with what we obtained
in the previous experiment, Table \ref{tab:puts:os}, we find higher values
for the location $\mu$ and scale $\sigma$ parameters, accounting for
the higher number of contributions per volunteer that has been
previously observed. However, the shape $\xi$ parameter is
substantially similar, with a {\em bell} shape leaning towards the
origin, indicating a pattern of a few users with many contributions
and many with few contributions. If we plot the model and the data
side by side, as shown in Figure \ref{fig:gev:w2} we see that the fit
is not so tight as in the previous case, with the model overestimating
the number of individuals with a high number contributions. This is
probably due to the fact that in the previous case, the user needed to
voluntarily reload the page to contribute the most generations, with a
{\em phase change} between the people that did so and the people that
just ran the experiment once until completion. In this case, however, the
user just needs to let it run, without that phase change and thus we obtain a
more {\em egalitarian} distribution of contributions, which rather
corresponds to a Weibull function, as shown in Figure
\ref{fig:ipstime:w2} (right hand side). The fitted parameters for this
distribution, shown in Table \ref{tab:puts:ww}, show a shape parameter
equal to 0.69, a value that is remarkably similar to the 0.5 value
found for time devoted to games in Chambers et
al. \cite{chambers2005measurement}. It is important to note that, in
this case, the number of PUTs made per client does not correspond to a
uniform amount of computation, since 100 generations with a random
population belong, in each case, to different number of
operations. In other words, the same number of PUTs will correspond to
different number of operations in each case. However, since clients
are also different and take a different time in each case, it is
difficult to ascertain how this might have an influence in the
statistical distribution.

>>>>

We are also interested in measuring the scaling properties of this
model, that is, the relationship between the number of IPs
participating in an experiment and the time it takes to complete
it. This is shown in Figure \ref{fig:ipstime:w2}, which displays every
experiment in terms of time (in milliseconds) vs. % (Paloma) Isn't "vs." informal?
the number of workers participating
in it. Although there is not a clear trend, the graph seems to say
that the time needed for finding the solution in this evolutionary
algorithm does not depend on the number of workers participating in
it. This does not mean that it is independent on the number of {\em
  simultaneous} users, which is probably the case although it is more
difficult to measure. There seems to be a trend towards a decreasing
standard deviation in the time, with more volunteers adding both {\em
  robustness} to an experiment, and certainty in the time it is taking. % (Paloma) Should "certainty" go in italics? Otherwise it's confusing. 
 However, more experiments would be needed to check this hypothesis.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.49\linewidth]{gev-fit-ww.png}
\includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
\caption{Ranked number of PUTs per worker and GEV fit (in light
  or blue color, left) 
and Weibull (in lighter or blue shade, right) for the {\sf NodIO-W$^2$} model.}  
\label{fig:gev:w2}
% File ../data/ips-puts.R
\end{figure}
%
As we did in the previous subsection, we have also fitted the number
of HTTP PUTs per worker to a  Generalized Extreme Value (GEV) function, with the result shown
in Table \ref{tab:puts:ww}. Comparing this table with what we obtained
in the previous experiment, Table \ref{tab:puts:os}, we find higher values
for the location $\mu$ and scale $\sigma$ parameters, accounting for
the higher number of contributions per volunteer that has been
previously observed. However, the shape $\xi$ parameter is
substantially similar, with a {\em bell} shape leaning towards the
origin, indicating a pattern of a few users with many contributions
and many with few contributions. If we plot the model and the data
side by side, as shown in Figure \ref{fig:gev:w2}, we see that the fit
is not so tight as in the previous case, with the model overestimating
the number of individuals with a high number contributions. This is
probably due to the fact that in the previous case the user needed to
voluntarily reload the page to contribute the most generations, with a
{\em phase change} between the people that did so and the people that
just ran the experiment once until completion. In this case, however, the
user just needs to let it run, without that phase change and thus we obtain a
more {\em egalitarian} distribution of contributions, which rather
corresponds to a Weibull function, as shown in Figure
\ref{fig:ipstime:w2} (right hand side). The fitted parameters for this
distribution, shown in Table \ref{tab:puts:ww}, show a shape parameter
equal to 0.69, a value that is remarkably similar to the 0.5 value
found for time devoted to games in Chambers et
al. \cite{chambers2005measurement}. It is important to note that, in
this case, the number of PUTs made per client does not correspond to a
uniform amount of computation, since 100 generations with a random
population belong, in each case, to different number of
operations. In other words, the same number of PUTs will correspond to
different number of operations in each case. However, since clients
are also different and take a different time in each case, it is
difficult to ascertain how this might have an influence in the
statistical distribution.
